Search Strategies for Functional Logic Programming

### 2012-01-13 papers/atps2012/paper.tex
Section on Curry started

\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}
\lstset{aboveskip=0.8ex,
        belowskip=0.8ex,
        xleftmargin=2ex,
        showstringspaces=false, % no special string space
        mathescape=true,
        flexiblecolumns=false,
        basewidth=0.52em,
        basicstyle=\small\ttfamily}
\lstset{literate={->}{{$\rightarrow{}\!\!\!$}}3
       }
\lstnewenvironment{curry}{}{}
\newcommand{\listline}{\vrule width0pt depth1.75ex}

@@

In the next section, we introduce the basics of Curry
that are necessary to understand the remaining part of the paper.
Different search strategies and their implementation
are discussed in Section~\ref{sec:strategies}.
These strategies are evaluated with a number of benchmarks
in Section~\ref{sec:benchmarks}
before we conclude in Section~\ref{sec:conclusions}

\section{Curry}

A short standard intro to Curry.

Mention that search strategy is not fixed
 (discuss top-level vs. encaps. search)

>>>>

\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}
\lstset{aboveskip=0.8ex,
        belowskip=0.8ex,
        xleftmargin=2ex,
        showstringspaces=false, % no special string space
        mathescape=true,
        flexiblecolumns=false,
        basewidth=0.52em,
        basicstyle=\small\ttfamily}
\lstset{literate={->}{{$\rightarrow{}\!\!\!$}}3
       }
\lstnewenvironment{curry}{}{}
\newcommand{\listline}{\vrule width0pt depth1.75ex}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\ccode}[1]{``\code{#1}''}

@@

In the next section, we briefly recall some principles
of functional logic programming and the programming language Curry
that are necessary to understand the remaining part of the paper.
Different search strategies and their implementation
are discussed in Section~\ref{sec:strategies}.
These strategies are evaluated with a number of benchmarks
in Section~\ref{sec:benchmarks}
before we conclude in Section~\ref{sec:conclusions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Functional Logic Programming and Curry}

Curry \cite{Hanus06Curry}
is a declarative multi-paradigm language combining features 
from functional programming (demand-driven evaluation, parametric
polymorphism, higher-order functions) and logic programming
(computing with partial information, unification, constraints).
Recent surveys are available in \cite{AntoyHanus10CACM,Hanus07ICLP}.
The syntax of Curry is close to Haskell \cite{PeytonJones03Haskell}.
Thus, variables and function names usually
start with lowercase letters and the names of type and data constructors
start with an uppercase letter. The application of $f$
to $e$ is denoted by juxtaposition (``$f~e$'').
In addition, Curry allows free (logic) 
variables in conditions and right-hand sides of defining rules.
The operational semantics is based on an optimal evaluation strategy
\cite{AntoyEchahedHanus00JACM} which is a conservative extension
of lazy functional programming and (concurrent) logic programming.

A Curry program consists of the definition of data types and
operations on these types. 
Note that in a functional logic language
operations might yield more than one result on the same input due to 
the logic programming features.
For instance, Curry contains a \emph{choice} operation defined by:
%
\begin{curry}
  x ? _ = x
  _ ? y = y
\end{curry}
%
Thus, the expression \ccode{0$~$?$~$1} has two values: \code{0} and \code{1}.
If expressions have more than one value, one wants to select
intended values according to some constraints,
typically in conditions of program rules.
A \emph{rule} has the form \ccode{$f~t_1\ldots{}t_n$ | $c$ = $e$}
where the (optional) condition $c$ is a \emph{constraint},
i.e., an expression of the built-in type
\code{Success}. For instance, the trivial constraint
\code{success} is a value of type \code{Success} that
denotes the always satisfiable constraint.
Thus, we say that a constraint $c$ is \emph{satisfied} if it
can be evaluated to \code{success}.
An \emph{equational constraint} $e_1 \,\code{=:=}\, e_2$ is satisfiable
if both sides $e_1$ and $e_2$ are reducible to unifiable values.
Furthermore, if $c_1$ and $c_2$ are constraints,
\code{$c_1\,$\&$\,\,c_2$} denotes their concurrent conjunction
(i.e., both argument constraints are concurrently evaluated).

As a simple example, consider the following Curry program
which defines a polymorphic 
data type for lists and operations to compute the 
concatenation of lists and the last element of a list:\footnote{Note
that lists are a built-in data type with a more convenient syntax, 
e.g., one can write \code{[x,y,z]} instead of \code{x:y:z:[]}
and \code{[a]} instead of the list type \ccode{List a}.}
%
\begin{curry}
  data List a = [] | a : List a    --[a] denotes "List a"

  -- "++" is a right-associative infix operator
  (++) :: [a] -> [a] -> [a]
  []     ++ ys = ys
  (x:xs) ++ ys = x : (xs ++ ys)

  last :: [a] -> a
  last xs | (ys ++ [z]) =:= xs
          = z                    where ys,z free
\end{curry}
%
Logic programming is supported by admitting function calls with free
variables (e.g., \code{(ys++[z])} in the rule defining \code{last})
and constraints in the condition of a defining rule. In contrast to
Prolog, free variables need to be declared explicitly to make their
scopes clear (e.g., \ccode{where ys,z free} in the example).
A conditional rule is applicable if its condition  is satisfiable.
Thus, the rule defining \code{last} states in its condition
that \code{z} is the last element of a given list
\code{xs} if there exists a list \code{ys}
such that the concatenation of \code{ys} and the one-element list
\code{[z]} is equal to the given list \code{xs}.

{\it
Mention that search strategy is not fixed
 (discuss top-level vs. encaps. search)}
